Combinatorial Proofs and Decomposition Theorems for First-order Logic\vspace{-3ex}

### 2021-01-25 main.tex
thm 32

x.(\cneg{p}x \vlor (\forall y.p(y)))}}{

@@

            \vlin{\vlor}{}{\sqn{\cneg{p}w, pz, \cneg{p}z \vlor (\forall y.py))}}{
              \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z, (\forall y.py))}}{

@@

\begin{proof}
  That $\graphof A$ is a logical cograph follows immediately from the definition and Theorem~\ref{thm:cograph}.
  The fact that every binder of $\graphof A$ is legal can be proved by
  structural induction on $A$.
  %% :
  %% \begin{itemize}
  %%   \item Trivial if $A$ is an atom.
  %%   \item The induction step uses the following fact:
  %%         if $M$ is a module of $\gG$ and $M' \subseteq M$, then $M'$ is a
  %%         module of $\gG$ if and only if it is a module of $\gG[X]$.
  %% \end{itemize}
  %% \todo{}\lutz{maybe mention also fn 2-5 in Dominic's arxiv-paper}
\end{proof}
%%\juihsuan{I'm still looking for a simple proof}
%
\begin{remark}
Note that $\graphof A$ is not necessarily a fograph if $A$ is not
rectified. If $A = (\forall x. px) \vlor (\forall x. qx)$, then
$\graphof A = \single x \ \single px \ \single x \ \single qx$,
the scope of each $x$-binder contains all the vertices, in particular,
the other $x$-binder. On the other hand, there are non-rectified
formulas which are translated to fographs by $\graphof\cdot$. For
example, in the graph of $(\exists x. px) \vlor (\exists x. qx)$,
both $x$-binders are legal, as they are not in each other's
scope:
$\edgepairpic{x}{px}\mkern9mu\edgepairpic{x}{qx}$.
\end{remark}
%
%% For this reason, we call a formula
%% \bfit{clean} if it does not contain subformulas of the form $(\forall
%% x.A)\vlor(\forall x.B)$ or $(\exists x.A)\vlan(\exists x.B)$,
%% \todo{need the four cases as in clear definition in fopws? Consider
%% $(\forall x.A)\vlor(B\vlor (\forall x.C))$. But then Lemma 6 doesn't hold.
%% (fopws proves only that there's a surjection from clear formulas onto fographs,
%% whose kernel is the six formula equalities)} and no
%% $x$-quantified formula occurs as subformula of another $x$-quantified
%% formula. Then we have:
%
%% \begin{lemma}
%%   If $A$ is clean iff\/ $\graphof A$ is a fograph.
%% \end{lemma}
%
%% \begin{proof}
%%   Induction on $A$, using  Theorem~\ref{thm:cograph}.
%% \end{proof}
%
%% Even though for every formula $A$ we can obtain an
%% equivalent clean formula by simply renaming some bound variables, this
%% is not unique up to $\alpha$-conversion, as is the case for
%% rectified formulas.
%
\noindent We define a congruence relation $\fequ$ on formulas by the following equations:
\begin{equation}
  \label{eq:fequ}
  \begin{array}{c@{\hskip1.5em}c}
    A \vlan B \fequ B \vlan A
    &
    (A \vlan B) \vlan C \fequ A \vlan (B \vlan C) 
    \\
    A \vlor B \fequ B \vlor A
    &
    (A \vlor B) \vlor C \fequ A \vlor (B \vlor C)
    \\
    \forall x . \forall y . A \fequ \forall y . \forall x . A
    &
    \forall x . (A \vlor B) \fequ (\forall x . A) \vlor B
    \\
    \exists x . \exists y . A \fequ \exists y . \exists x . A
    &
    \exists x . (A \vlan B) \fequ (\exists x . A) \vlan B
  \end{array}
\end{equation}
where $x$ must not be free in $B$ in the last two equations.
Two formulas $A$ and $B$ are \bfit{equivalent} if $A \fequ B$. The following
theorem shows that the set of fographs can been seen as the quotient
$\FORM/\mathord\fequ$.
\begin{thm}
  Let $A,B$ be rectified formulas. Then
  $$
  A\fequ B \iff \graphof A =\graphof B
  $$
\end{thm}

@@

For proving this theorem, we have to show that no inference rule in
$\FOMLS$ can introduce a bimatching. To simplify the argument, we
introduce the \bfit{frame}~\cite{hughes:unifn} of the fograph $\gC$,
which is a linked (proposional) cograph in which the dependencies
between the binders in $\gC$ are encoded as links.

@@

\begin{proof}[Proof of Theorem~\ref{thm:MLS1->fonet}]
  From $\Deri$ we construct a derivation $\frameof\Deri$ of $\frameof
  A$ in the propositional fragment of $\FOMLS$, such that
  \hbox{$\fographof{\frameof\Deri}=\frameof{\fographof\Deri}$}.  The
  rules $\aiD,\tttD,\mix$ and $\sw$ are translated trivially, and for
  $\fequ$, it suffices to observe that the frame construction is
  invariant under $\fequ$. Finally, for the rules $\forall$ and
  $\exists$, proceed as follows. Every instance of $\forall$ is
  replaced by the derivation on the right below:\footnote{For better
  readability we omit superfluous parentheses, knowing that we always
  have $\fequ$ incorporating associativity and commutativity of
  $\vlan$ and $\vlor$.} \vadjust{\vskip-2ex}
  \begin{equation*}
    \scalebox{.95}{$
    \vlinf{\forall}{}{\Scons{\forall y_j.\ttt}}{\Scons{\ttt}}
    %\qquad
    \;\leadsto
    %\qquad
    \vlderivation{
      \vlde{\set{\sw,\fequ}}{\DDeri_2}{\Scons{q_{h_1,j}\vlor\cdots\vlor
          q_{h_j,j}\vlor(\dual q_{h_1,j}\vlan\cdots\vlan\dual
          q_{h_j,j}\vlan\ttt)}}{
        \vlde{\set{\aiD,\tttD}}{\DDeri_1}{
          \Scons{(q_{h_1,j}\vlor \dual q_{h_1,j})\vlan\cdots\vlan(q_{h_j,j}\vlor \dual q_{h_j,j})\vlan\ttt}}{
          \vlhy{\ttt}}}}
    $}
  \end{equation*}
  where $h_1,\ldots,h_j$ range over the indices of the existential
  binders dependend on that $y_j$. It is easy to see how $\DDeri_1$ is
  constructed, and for $\DDeri_2$ see,
  e.g. \cite{ATS:esslli2019,gug:str:01,brunnler:tiu:01,dissvonlutz}. Then, every
  occurrence of $\forall y_j.F$ is replaced by
  $q_{h_1,j}\vlor\cdots\vlor q_{h_j,j}\vlor (\dual
  q_{h_1,j}\vlan\cdots\vlan\dual q_{h_j,j}\vlan F)$ in the derivation
  below that $\forall$-instance. Now, observe that all instances of
  the $\exists$-rule introducing $x_i$ dependend on $y_j$ must occur
  below in the derivation (otherwise $\Deri$ would not be rectified).
  Now consider such an instance $\vlinf{\exists}{}{\Scons{\exists
      x_i.B}}{\Scons{B\ssubst {x_i} t}}$.  Its context $\Sconhole$
  must contain all the $\forall y_j$ the $\exists x_i$ depends on,
  such that $B$ is in their scope. Following the translation of the
  $\forall$ rules above, we can therefore translate the $\exists$-rule
  instance by the following derivation
  \begin{equation*}
    \vlder{\set{\sw,\fequ}}{\DDeri_3}{ S_0\cons{S_1\cons{\cdots
          S_{k_i-1}\cons{S_{k_i}\cons{\dual q_{i,k_1}\vlan\dual
              q_{i,k_2}\vlan\cdots q_{i,k_i}\vlan B'}}\cdots}}}{
      S_0\cons{\dual q_{i,k_1}\vlan S_1\cons{\dual
          q_{i,k_2}\vlan\cdots S_{k_i-1}\cons{\dual q_{i,k_i}\vlan
            S_{k_i}\cons{B'}}\cdots}}}
  \end{equation*}
  where $k_1,\ldots,k_i$ are the indices of of the universal binders
  on which that $x_i$ depends, and $B'$ is $B$ in which all predicates
  are replaced by a nullary one (step~3 in the frame construction). The
  derivation $\DDeri_3$ can be constructed in the same way as
  $\DDeri_2$.

@@

\begin{proof}
   By \cite[Proposition~7.6.1]{str:07:RTA}, we have a derivation
   $\vlder{\set{\wrD,\PE\fequ}}{\DDeri}{\PE H}{\PE G}$. In order to
   lift $\DDeri$, we might need to reorganize the instancese of
   $\wD$. If $H$ contains a subformula $\forall x.A$ which is not
   present in $G$, the $\wD$-instances in $\DDeri$ could introduce the
   parts of the propositional encoding $x\vlor A$ independently.  We
   say that an instance $\rr_1$ of $\wD$ in $\Deri$ is \emph{in the
   scope} of an instance $\rr_2$ of $\wD$ if $\rr_1$ introduced
   formulas that contain a free variable $x$ (i.e., $x$ occurs in a
   term in a predicate) and $\rr_2$ introduces the atom $x$ as a
   subformula (i.e. the propositional encoding of the binder $x$).  We
   can now permute the $\wD$-instances in $\DDeri$ such that whenever
   a rule instance $\rr_1$ depends on an instance $\rr_2$, then
   $\rr_2$ occurs below $\rr_1$ in $\DDeri$.
   We can now lift $\DDeri$ stepwise. First, observe that each line of
   in $\DDeri$ is $\PE\fequ$-equivalent to the propositional encoding $\PE P$ of a
   first-order formula $P$.  We now have to show
   that each other instance of $\wrD$ in $\DDeri$ is indeed the image of a correct
   application of $\wD$ or $\wfaD$ in first-order logic.
   If we have a $\wD$
   \begin{equation*}
     \vlinf{\wD}{}{\PE S\cons{x\vlor \PE A}}{\PE S\cons{\PE A}}
     \qquor
     \vlinf{\wD}{}{\PE S\cons{(x\vlor \PE B)\vlor \PE A}}{\PE S\cons{\PE A}}
   \end{equation*}
   then $x$ cannot occur freely in $A$, as otherwise the fibration
   property would be violated. We can therefore lift these instances to
   \begin{equation*}
     \vlinf{\wfaD}{}{S\cons{\forall x. A}}{S\cons{A}}
     \qquor
     \vlinf{\wD}{}{S\cons{(\forall x. B)\vlor A}}{S\cons{A}}
   \end{equation*}
   respectively.     
  %% A subformula of $\PE B$ is called \emph{weak} if it has been
  %% introduced by a weakening. All subformulas of a weak formula are
  %% also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  %% \emph{weak pair} if $\PE B\fequ\Scons{B'\vlor B''}$ for some context
  %% $\Sconhole$. We can assume without loss of generality that
  %% whenever weak subformulas $B'$ and $B''$ form a weak pair, they have
  %% been introduced by the same instance of $\wrD$ in
  %% $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  %% brought into this form by simple rule permutations, and then
  %% collapsing several weakenings into a single one.}  Now we show that
  %% $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  %% weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  %% delete all atoms in the scope of the corresponding quantifier,
  %% because $\phi'$ is a fibration on the binding graph. Hence, each line
  If a  weakening happens inside a subformula $x\vlor \PE C$ or $x\vlan
  \PE C$ in $\DDeri$, then there are the following cases:
  \begin{equation*}%\small
    \scalebox{.9}{$
    \vlinf{\wD}{}{\PE S\cons{x\vlor\PE  D\vlor \PE C}}{\PE S\cons{x\vlor\PE  C}}
    \quad
    \vlinf{\wD}{}{\PE S\cons{x\vlan (\PE D\vlor \PE C)}}{\PE S\cons{x\vlan\PE  C}}
    \quad
    \vlinf{\wD}{}{\PE S\cons{(x\vlor\PE  D)\vlan \PE C}}{\PE S\cons{x\vlan \PE C}}
    $}
  \end{equation*}
  The first two cases can be lifted to
  \begin{equation*}
    \vlinf{\wD}{}{S\cons{\forall x.( D\vlor C)}}{S\cons{\forall x.  C}}
    \quand
     \vlinf{\wD}{}{S\cons{\exists x.( D\vlor C)}}{S\cons{\exists x.  C}}
  \end{equation*}
  respectively. But in the third case,  an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur. All other situations can be lifted trivially, giving us
  $\vlder{\set{\wrD,\wfaD\fequ}}{\Deri}{H}{G}$ as desired.
\end{proof}

@@

We first study the interactions between two rules. Certain cases are unsolved at
this stage, and they are considered later when we study the interactions between
two non-$\fequ$ rule instances seperated by $\fequ$. Only non-trivial cases are
considered here:
\begin{itemize}
\item $\rr_1/\rr_2$, where $\rr_1 \in \set{\wrD, \wfaD}$ and $\rr_2 \in
\set{\acD, \cfaD, \me, \mfaD, \mexD}$:

>>>>

@@

            \vlin{\vlor}{}{\sqn{\cneg{p}w, pz, \cneg{p}z \vlor (\forall y.py)}}{
              \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z, \forall y.py}}{

@@

\begin{proof}
  That $\graphof A$ is a logical cograph follows immediately from the definition and Theorem~\ref{thm:cograph}.
  The fact that every binder of $\graphof A$ is legal can be proved by
  structural induction on $A$.
  %% :
  %% \begin{itemize}
  %%   \item Trivial if $A$ is an atom.
  %%   \item The induction step uses the following fact:
  %%         if $M$ is a module of $\gG$ and $M' \subseteq M$, then $M'$ is a
  %%         module of $\gG$ if and only if it is a module of $\gG[X]$.
  %% \end{itemize}
  %% \todo{}\lutz{maybe mention also fn 2-5 in Dominic's arxiv-paper}
\end{proof}
%%\juihsuan{I'm still looking for a simple proof}
%
\begin{remark}
Note that $\graphof A$ is not necessarily a fograph if $A$ is not
rectified. If $A = (\forall x. px) \vlor (\forall x. qx)$, then
$\graphof A = \single x \ \single px \ \single x \ \single qx$,
the scope of each $x$-binder contains all the vertices, in particular,
the other $x$-binder. On the other hand, there are non-rectified
formulas which are translated to fographs by $\graphof\cdot$. For
example, in the graph of $(\exists x. px) \vlor (\exists x. qx)$,
both $x$-binders are legal, as they are not in each other's
scope:
$\edgepairpic{x}{px}\mkern9mu\edgepairpic{x}{qx}$.
\end{remark}
%
%% For this reason, we call a formula
%% \bfit{clean} if it does not contain subformulas of the form $(\forall
%% x.A)\vlor(\forall x.B)$ or $(\exists x.A)\vlan(\exists x.B)$,
%% \todo{need the four cases as in clear definition in fopws? Consider
%% $(\forall x.A)\vlor(B\vlor (\forall x.C))$. But then Lemma 6 doesn't hold.
%% (fopws proves only that there's a surjection from clear formulas onto fographs,
%% whose kernel is the six formula equalities)} and no
%% $x$-quantified formula occurs as subformula of another $x$-quantified
%% formula. Then we have:
%
%% \begin{lemma}
%%   If $A$ is clean iff\/ $\graphof A$ is a fograph.
%% \end{lemma}
%
%% \begin{proof}
%%   Induction on $A$, using  Theorem~\ref{thm:cograph}.
%% \end{proof}
%
%% Even though for every formula $A$ we can obtain an
%% equivalent clean formula by simply renaming some bound variables, this
%% is not unique up to $\alpha$-conversion, as is the case for
%% rectified formulas.
%
\noindent We define a congruence relation $\fequ$ on formulas by the following equations:
\begin{equation}
  \label{eq:fequ}
  \begin{array}{c@{\hskip1.5em}c}
    A \vlan B \fequ B \vlan A
    &
    (A \vlan B) \vlan C \fequ A \vlan (B \vlan C) 
    \\
    A \vlor B \fequ B \vlor A
    &
    (A \vlor B) \vlor C \fequ A \vlor (B \vlor C)
    \\
    \forall x . \forall y . A \fequ \forall y . \forall x . A
    &
    \forall x . (A \vlor B) \fequ (\forall x . A) \vlor B
    \\
    \exists x . \exists y . A \fequ \exists y . \exists x . A
    &
    \exists x . (A \vlan B) \fequ (\exists x . A) \vlan B
  \end{array}
\end{equation}
where $x$ must not be free in $B$ in the last two equations.
Two formulas $A$ and $B$ are \bfit{equivalent} if $A \fequ B$. We have the
following theorem:
\begin{thm}
  Let $A,B$ be rectified formulas. Then
  $$
  A\fequ B \iff \graphof A =\graphof B
  $$
\end{thm}

@@

For proving this theorem, we have to show that no inference rule in
$\FOMLS$ can introduce a bimatching. To simplify the argument, we
introduce the \bfit{frame}~\cite{hughes:unifn} of the linked fograph $\gC$,
which is a linked (proposional) cograph in which the dependencies
between the binders in $\gC$ are encoded as links.

@@

\begin{proof}[Proof of Theorem~\ref{thm:MLS1->fonet}]
  From $\Deri$ we construct a derivation $\frameof\Deri$ of $\frameof
  A$ in the propositional fragment of $\FOMLS$, such that
  \hbox{$\fographof{\frameof\Deri}=\frameof{\fographof\Deri}$}.  The
  rules $\aiD,\tttD,\mix$ and $\sw$ are translated trivially, and for
  $\fequ$, it suffices to observe that the frame construction is
  invariant under $\fequ$. Finally, for the rules $\forall$ and
  $\exists$, proceed as follows. Every instance of $\forall$ is
  replaced by the derivation on the right below:\footnote{For better
  readability we omit superfluous parentheses, knowing that we always
  have $\fequ$ incorporating associativity and commutativity of
  $\vlan$ and $\vlor$.} \vadjust{\vskip-2ex}
  \begin{equation*}
    \scalebox{.95}{$
    \vlinf{\forall}{}{\Scons{\forall y_j.\ttt}}{\Scons{\ttt}}
    %\qquad
    \;\leadsto
    %\qquad
    \vlderivation{
      \vlde{\set{\sw,\fequ}}{\DDeri_2}{\Scons{q_{h_1,j}\vlor\cdots\vlor
          q_{h_n,j}\vlor(\dual q_{h_1,j}\vlan\cdots\vlan\dual
          q_{h_n,j}\vlan\ttt)}}{
        \vlde{\set{\aiD,\tttD}}{\DDeri_1}{
          \Scons{(q_{h_1,j}\vlor \dual q_{h_1,j})\vlan\cdots\vlan(q_{h_n,j}\vlor \dual q_{h_n,j})\vlan\ttt}}{
          \vlhy{\ttt}}}}
    $}
  \end{equation*}
  where $h_1,\ldots,h_n$ range over the indices of the existential
  binders dependend on that $y_j$. It is easy to see how $\DDeri_1$ is
  constructed, and for $\DDeri_2$ see,
  e.g. \cite{ATS:esslli2019,gug:str:01,brunnler:tiu:01,dissvonlutz}. Then, every
  occurrence of $\forall y_j.F$ is replaced by
  $q_{h_1,j}\vlor\cdots\vlor q_{h_j,j}\vlor (\dual
  q_{h_1,j}\vlan\cdots\vlan\dual q_{h_j,j}\vlan F)$ in the derivation
  below that $\forall$-instance. Now, observe that all instances of
  the $\exists$-rule introducing $x_i$ dependend on $y_j$ must occur
  below in the derivation (otherwise $\Deri$ would not be rectified).
  Now consider such an instance $\vlinf{\exists}{}{\Scons{\exists
      x_i.B}}{\Scons{B\ssubst {x_i} t}}$.  Its context $\Sconhole$
  must contain all the $\forall y_j$ the $\exists x_i$ depends on,
  such that $B$ is in their scope. Following the translation of the
  $\forall$ rules above, we can therefore translate the $\exists$-rule
  instance by the following derivation
  \begin{equation*}
    \vlder{\set{\sw,\fequ}}{\DDeri_3}{ S_0\cons{S_1\cons{\cdots
          S_{k_l-1}\cons{S_{k_l}\cons{\dual q_{i,k_1}\vlan\dual
              q_{i,k_2}\vlan\cdots q_{i,k_l}\vlan B'}}\cdots}}}{
      S_0\cons{\dual q_{i,k_1}\vlan S_1\cons{\dual
          q_{i,k_2}\vlan\cdots S_{k_l-1}\cons{\dual q_{i,k_k}\vlan
            S_{k_l}\cons{B'}}\cdots}}}
  \end{equation*}
  where $k_1,\ldots,k_l$ are the indices of of the universal binders
  on which that $x_i$ depends, and $B'$ is $B$ in which all predicates
  are replaced by a nullary one (step~3 in the frame construction). The
  derivation $\DDeri_3$ can be constructed in the same way as
  $\DDeri_2$.

@@

\begin{proof}
   By \cite[Proposition~7.6.1]{str:07:RTA}, we have a derivation
   $\vlder{\set{\wrD,\PE\fequ}}{\DDeri}{\PE H}{\PE G}$. In order to
   lift $\DDeri$, we might need to reorganize the instancese of
   $\wD$. If $H$ contains a subformula $\forall x.A$ which is not
   present in $G$, the $\wD$-instances in $\DDeri$ could introduce the
   parts of the propositional encoding $x\vlor A$ independently.  We
   say that an instance $\rr_1$ of $\wD$ in $\Deri$ is \emph{in the
   scope} of an instance $\rr_2$ of $\wD$ if $\rr_1$ introduced
   formulas that contain a free variable $x$ (i.e., $x$ occurs in a
   term in a predicate) and $\rr_2$ introduces the atom $x$ as a
   subformula (i.e. the propositional encoding of the binder $x$).  We
   can now permute the $\wD$-instances in $\DDeri$ such that whenever
   a rule instance $\rr_1$ depends on an instance $\rr_2$, then
   $\rr_2$ occurs below $\rr_1$ in $\DDeri$.
   We can now lift $\DDeri$ stepwise. First, observe that each line of
   in $\DDeri$ is $\PE\fequ$-equivalent to the propositional encoding $\PE P$ of a
   first-order formula $P$.  We now have to show
   that each other instance of $\wrD$ in $\DDeri$ is indeed the image of a correct
   application of $\wD$ or $\wfaD$ in first-order logic.
   If we have a $\wD$
   \begin{equation*}
     \vlinf{\wD}{}{\PE S\cons{x\vlor \PE A}}{\PE S\cons{\PE A}}
     \qquor
     \vlinf{\wD}{}{\PE S\cons{(x\vlor \PE B)\vlor \PE A}}{\PE S\cons{\PE A}}
   \end{equation*}
   then $x$ cannot occur freely in $A$, as otherwise the fibration
   property would be violated. We can therefore lift these instances to
   \begin{equation*}
     \vlinf{\wfaD}{}{S\cons{\forall x. A}}{S\cons{A}}
     \qquor
     \vlinf{\wD}{}{S\cons{(\forall x. B)\vlor A}}{S\cons{A}}
   \end{equation*}
   respectively.     
  %% A subformula of $\PE B$ is called \emph{weak} if it has been
  %% introduced by a weakening. All subformulas of a weak formula are
  %% also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  %% \emph{weak pair} if $\PE B\fequ\Scons{B'\vlor B''}$ for some context
  %% $\Sconhole$. We can assume without loss of generality that
  %% whenever weak subformulas $B'$ and $B''$ form a weak pair, they have
  %% been introduced by the same instance of $\wrD$ in
  %% $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  %% brought into this form by simple rule permutations, and then
  %% collapsing several weakenings into a single one.}  Now we show that
  %% $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  %% weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  %% delete all atoms in the scope of the corresponding quantifier,
  %% because $\phi'$ is a fibration on the binding graph. Hence, each line
  If a  weakening happens inside a subformula $x\vlor \PE C$ or $x\vlan
  \PE C$ in $\DDeri$, then there are the following cases:
  \begin{equation*}%\small
    \scalebox{.9}{$
    \vlinf{\wD}{}{\PE S\cons{x\vlor\PE  D\vlor \PE C}}{\PE S\cons{x\vlor\PE  C}}
    \quad
    \vlinf{\wD}{}{\PE S\cons{x\vlan (\PE D\vlor \PE C)}}{\PE S\cons{x\vlan\PE  C}}
    \quad
    \vlinf{\wD}{}{\PE S\cons{(x\vlor\PE  D)\vlan \PE C}}{\PE S\cons{x\vlan \PE C}}
    $}
  \end{equation*}
  The first two cases can be lifted to
  \begin{equation*}
    \vlinf{\wD}{}{S\cons{\forall x.( D\vlor C)}}{S\cons{\forall x.  C}}
    \quand
     \vlinf{\wD}{}{S\cons{\exists x.( D\vlor C)}}{S\cons{\exists x.  C}}
  \end{equation*}
  respectively. But in the third case,  an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur. All other situations can be lifted trivially, giving us
  $\vlder{\set{\wrD,\wfaD,\fequ}}{\Deri}{H}{G}$ as desired.
\end{proof}

@@

We first study the interactions between two rules. Certain cases are unsolved at
this stage, and they are considered later when we study the interactions between
two non-$\fequ$ rule instances seperated by $\fequ$. Only non-trivial cases are
presented here:
\begin{itemize}
\item $\rr_1/\rr_2$, where $\rr_1 \in \set{\wrD, \wfaD}$ and $\rr_2 \in
\set{\acD, \cfaD, \me, \mfaD, \mexD}$:
